Modular Inverse Reinforcement Learning on Human Motion

### 2015-02-11 paper.tex
set up for modification

Human is able to learn accomplishing complicated tasks much faster than machines
can do. However, most of the tasks can be decomposed into subtasks. A human may
already have the capacity to accomplish the subtasks. He simply transfers the
knowledge of the subtasks and integrate them for the objective task. For example,
when a human learns to drive, he has much prior domain knowledge to help him
with this objective task. He integrates his skills of controlling the velocity,
avoiding other vehicles, navigating, responding to traffic lights, and so on.

Similar ideas are also adopted in the learning literature. Reinforcement learning
suffers from the curse of dimensionality. It is inefficient for a learning agent
to learn a complicated task from scratch. So learning algorithms including
hierarchical reinforcement learning \cite{dietterich2000hierarchical}, modular
reinforcement learning \cite{sprague2003multiple} are proposed. They also
decompose the objective task into subtasks. The learning algorithms first learn
the subtasks independently, then learn how to combine these subtasks.

A natural question to ask is how human integrates skills for subtasks, and
whether we can use the same method for integrating subtasks in reinforcement
learning problems. In this paper, we analyze human's behavior of
accomplishing a task composite of various subtasks. We collected human motion
data and try to interpret the behavior using a way of
inverse reinforcement learning. With the best of our knowledge, this approach is
novel in the literature.

@@

\section{Multi-objective Sidewalk Domain}
\label{sec:domain}

@@

Figure~\ref{fig:avatar} shows the domain that we use in this paper. The red
cubes represent obstacles, that the player would be punished if running
into. The blue balls represent targets, that the player would be rewarded if
collected. There is also a gray path on the ground that the player can follow.
Naturally, this domain has three subtasks, 1) following the path, 2)
collecting targets, 3) avoiding obstacles.
This is an experiment used in the
literature to evaluate modular reinforcement learning
\cite{rothkopf2013modular}.
We ask our volunteers to have trackers and virtual reality displayers equipped. He
can see the envrionment through the displayer in front of his eyes. So he can
walk as if he is walking in the virtual domain.

@@

We need some basic concepts in reinforcement learning to proceed. For a state, action
pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
state $s$. The {\em policy} of a state is the action with the maximum Q
value \cite{rl}. So how to we determine the global policy from the modules, or
subtasks? In the literature, work has been done to integrate the Q function
\cite{koller1999computing} or compromise on policies directly
\cite{thomas2012motor}.

@@

We train the modules first to get $Q_i$ before running the inverse reinforcement
learning algorithm. For each module, the agent only considers the closest target
and the closest obstacle. For the path module, the path is defined as segments
of waypoints, so the closest waypoint is considered. The agent takes the
distance and angle to the closest objects as the state representation.

To make our weights represent the significance of the modules, we
normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
is 1 for collecting a target, -1 for running into an obstacle. We define the
value function directly for the path module to have a path following
performance, as it is tricky to give reward for such performance.

>>>>

A large body of evidence from human and other primate studies studies 
suggests that they use reinforcement signals to learn tasks and reinforcement 
learning has proved to be a good model for this process.
However humans are able to learn and carry out very complex tasks involving 
many different objectives.

How can they do this? The likelihood is that they use rinforcement in this 
comtext as well but most formal reinforcement learning models do not scale 
well with increasing complexity.

One promising possibiity is that the complex task can be broken down into 
subtasks that are each learned separately \cite{sprague2003multiple,
rothkopf2013modular}. 
For certain task venues, this decomposition allows the behavior in the complex 
task to be specified as a eighted sum of individal sub-task priorities.
This paper explores the use of such a decomposition to characterize complex 
behavior of subjects in a structure virtual reality environment. The 
methodology of \cite{rothkopf2013modular} allows the task priorities to be
estimated using the subjects' behavioral data.

Our experimental analysis shows that the modular reinforcement can be a 
surprisingly good model of behavior, predicting individual subjects' sub-task 
priorities in a way that explains their beahvioral choices.

@@

\section{Multi-objective Domain}
\label{sec:domain}

@@

%FIXME
In Figure~\ref{fig:avatar} shows the domain that we use in this paper. The red
cubes represent obstacles, that the player would be punished if running
into. The blue balls represent targets, that the player would be rewarded if
collected. There is also a gray path on the ground that the player can follow.
Naturally, this domain has three subtasks, 1) following the path, 2)
collecting targets, 3) avoiding obstacles.
This is an experiment used in the
literature to evaluate modular reinforcement learning
\cite{rothkopf2013modular}.
We ask our volunteers to have trackers and virtual reality displayers equipped. He
can see the envrionment through the displayer in front of his eyes. So he can
walk as if he is walking in the virtual domain.

@@

%FIXME
For a state, action
pair, $s$ and $a$, $Q(s, a)$ evaluates the utility of taking action $a$ from
state $s$. The {\em policy} of a state is the action with the maximum Q
value \cite{rl}. So how to we determine the global policy from the modules, or
subtasks? In the literature, work has been done to integrate the Q function
\cite{koller1999computing} or compromise on policies directly
\cite{thomas2012motor}.

@@

We trained the modules first to get $Q_i$ before running the inverse reinforcement
learning algorithm. For each module, the agent only considers the closest target
and the closest obstacle. For the path module, the path is defined as segments
of waypoints, so the closest waypoint is considered. The agent takes the
distance and angle to the closest objects as the state representation.

To make our weights represent the significance of the modules, we
normalize the sub-MDPs with the unit (positive or negative) rewards. The reward
is 1 for collecting a target, -1 for running into an obstacle. We define the
value function directly for the path module to have a path following
performance, as it is tricky to give reward for such performance.
% emphasize that path following is not an RL module?
